%!TEX root = ../dissertation.tex
\chapter{Data analysis for the experiments}
\label{AppendixB}

\section{Quantum thermalization}
For all entanglement entropy data we average over both copies after the beam splitter operation. Additionally, we take the mean of equivalent symmetric subsystems in each copy for entanglement entropy dynamics (fig.~\ref{fig:EEDyn}), for the volume law data (fig.~\ref{fig:volume}) we average over all contiguous subsystems of the given size. Finally, for the data in fig.~\ref{fig:MI}, we average over all contiguous and non-contiguous subsystems of the given volume, because the mutual information can sample correlations between non-contiguous subsystems.  For all of the entanglement entropy data, the error bars are S.E.M. For all of the number counting, the error bars are S.E.M. except when stated otherwise.

\section{Many-body localization}
For all data in fig.~\ref{fig:MBL_ETH_breakdown}~,~\ref{fig:MBL_localization} we use 197 unique disorder patterns and preform a running average over them by randomly sampling a given number of realizations and treating them as independent measurements of the same system. Additionally in fig.~\ref{fig:MBL_localization} the data is averaged over 6 middle sites of the chain. For fig.~\ref{fig:MBL_entropy_dynamics}~,~\ref{fig:MBL_scalings} we use 4 different disorder patterns. We first average each observable over different outcomes of the same disorder and subsequently perform the average over different disorder realizations. 

\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline Figure &disorders& post selected shots/point\\
		\hline \ref{fig:MBL_ETH_breakdown}A: W=1J& 197 & 64, 82, 68, 63, 58, 55, 133, 142, 192, 180, 160, 108\\ 
		\hline \ref{fig:MBL_ETH_breakdown}A: W=8.9J& 197 & 134, 135, 143, 140, 156, 152, 144, 126, 123, 128, 126, 147 \\ 
		\hline \ref{fig:MBL_ETH_breakdown}B & 197 &106, 113, 118, 129, 191, 186, 101, 186, 185, 104, 110, 124, 102\\ 
		\hline \ref{fig:MBL_localization}A: W=1J& 197 & 160\\ 
		\hline \ref{fig:MBL_localization}A: W=8.9J& 197 & 110 \\ 
		\hline \ref{fig:MBL_localization}B & 197 & 129, 191, 101, 186, 185, 104, 110, 124, 102 \\ 
		\hline \ref{fig:MBL_entropy_dynamics} & 4 & Avg/Disorder:210.5, 305.75, 264.25, 219, 263.5, 196.25\\
		\hline \ref{fig:MBL_scalings} & 4 & Avg/Disorder:740.25\\
		\hline
	\end{tabular} 
\end{center}