%!TEX root = ../dissertation.tex
\chapter{Long coherent dynamics in optical lattices}

\section{Law entropy initial states}
Quantum mechanics gives us tools to study highly correlated system. The most famous example is an entangled Bell state \cite{something}, which can be written in context of two spin $\frac{1}{2}$ particles as 
\begin{equation}
\left| \psi \right>=(\left| \uparrow \downarrow \right>+ \left| \downarrow \uparrow \right>)/\sqrt{2},
\end{equation}
where up and down arrows represent one of the two states of each spin. The remarkable property of this state is, that no matter which basis one would chose to measure both particles in, the results of the measurement will always be anti correlated. This entanglement could further be utilized as a resource for variety of application ranging form metrology and sensing \cite{sombody} to quantum computing \cite{Chuang book} and secure quantum communications \cite{something}. However those correlations are very fragile: addition of any amount of classical mixture of $\left| \uparrow \right>$ and $\left| \downarrow \right>$ states to any of the spins immediately destroys the perfect correlations between the two, that ultimately limits the practical applications of quantum techniques. In this chapter we will show, how using the single site control over the atoms in our quantum gas microscope, we can prepare and control highly entangled many-particle quantum systems, in particular in the context of quantum simulations. 

In our experiments the goal is to study the dynamics of strongly correlated many-body systems. Due to the probabilistic nature of quantum mechanics, in order to learn information about a quantum state, one needs to have an access to multiple copies of the same state. Then by doing repeated measurements, the probability distribution of occupation of each particular basis state can be obtained. This means, that we also need a way to deterministically prepare the same initial state with high fidelity. 

In order to create an entangled state, one starts with a product state and then apply an appropriate entangling Hamiltonian to arrive at desired state.  For atoms in the optical lattices one of the examples of a conceptually simple initial state is a product state of given atom numbers on individual lattice sites. For all the experiments described bellow, we would like to prepare strings of desired length with a single atom on each site. There are to ways to achieve that: one is a bottom up approach, where by using single site addressing and control one can trap a single atom into a tightly focused optical tweezer. Then by using Raman cooling techniques the atom's vibrational state can be cooled to almost perfect ground state in all three dimensions \cite{adamo, selim}. Finally, by subsequent rearranging an array of such tweezer traps, in order to eliminate defects, one can achieve deterministic strings of up to 51 atoms \cite{misha} or $?\times?$ two dimensional arrays with unity feeling \cite{broweys}. This platform has successfully enabled study of spin models, using Rydberg blockade techniques \cite{misha}. However, the current state of the art cooling fidelities do not allow for efficient preparation of single band Hubbard model with large number of particles, due to residual exited state fraction scaling exponentially with the number of atoms.

An alternative route is a top down approach, where one uses an ensemble of atoms in order to first crate a macroscopic number of atoms in the absolute ground state \cite{BEC, DFG}, then selectively isolate those, and finally adiabatically load them into an optical lattice \cite{Greiner2002}. Then, by increasing the onsight interaction strength between the atoms, by ramping up the lattice depth, one arrives at the Mott insulating state (see fig.~\ref{fig:CTE_MI}A), where each lattice site is occupied by an integer number of atoms according to a local chemical potential, with a vanishing atom number variance \cite{Bakr2010, Bloch MI}. Due to the finite ramp time between two many-body states it is impossible to remain exactly in the ground state of the system \cite{subir phase transition}. Therefor the resulting atom number distribution deviates from the ideal one by "smearing" the boundary between the shells with different occupation numbers (see fig~\ref{CTE_MI}B). However, if the Mott insulator is sufficiently large we can find a region with unity filling up to $12$ sites long. Hence, to achieve our desired target state, we just need to isolate such a region, and ensure that it doesn't get disturbed by the adjacent atoms during subsequent evolution.

\section{Cutting procedure}

In order to isolate a desired region of the lattice we preform so called "cutting" procedure. Using the DMD we project finite size lattice potential on top of initial $2\mathrm{D}$ lattice, in which the Mott insulator was originally crated. Then by applying slow varying anticonfinement beam and switching off the the $2\mathrm{D}$ lattice we make the atoms leave, except for those that were kept by the DMD potential (see fig.~\ref{fig:CTE_cutting}A). Finally we can reload the atom back into the $2\mathrm{D}$ lattice by ramping it back up and switching off the DMD potential. By repeating this procedure along both directions, we managed to isolate plaquettes up to $12\times2$ in size (see fig.~\ref{fig:CTE_cutting}B). This technique allows us to achieve $99\%$ single atom transfer fidelity form initial Mott insulator. Cutting more then $12$ sites along one direction using this method is limited by the available power of the DMD beam. Similar results have been achieved using state dependent light shifts and microwave pulses \cite{Bloch single site addresing}.

By increasing the DMD beam power we could in principal initialize even longer chains, however there is another difficulty in our system. In order to achieve high fidelity of $n=1$ shell we rely on the idea of entropy redistribution in our system \cite{some entropy redistribution}. It can be summarized in the following way: crossing the transition from Superfluid to Mott insulator inject the system with certain amount of excitations. If the system would be homogeneous those excitations would spread evenly across the system. However, there is a strong inhomogeneity in the regions where shells of different occupation number touch one another. Since the density has to smoothly connect from one occupation to another, it deviates from an integer value creating regions of suprefluid. Superfluid has significantly smaller gap between law lying states compared to a Mott shell, hence it is energetically favorable for the excitations to concentrate in the regions between the shells. From the above considerations, it is clear that this mechanism works better for larger ration of the superfluid region to the Mott shells. This technique has also found a great success in Fermi gas microscopes, where manual increase of the entropy reservoir region led to significant entropy reductions in the gapped part of the system \cite{Mazurenko2017, Chiu2018}.

 From the analysis above it fallows, that in order to create $n=1$ shell with high fidelity, it is advantageous to have it be surrounded by $n=0$ and $n=2$ shells. However, due to large amount of disorder in our lattice it is particularly challenging to create large, round Mott insulators, such that one can isolate a single line at unity filling longer then $\sim 16$ sites. This means that, in order to crate longer chains one has to switch to a different Mott insulator geometry to begin with. One simple solution that comes to mind is to make a rectangular box confinement instead of a harmonic one. In order to have well defined shells in this case one can use a magnetic field gradient in order to create a uniform tilt. We realize such configuration in our system in a $24$ site long box (see fig.~\ref{fig:CTE_MI_box}), then by using a cutting procedure along the other direction we were able to achieve $24\times 2$ plaquettes. In this case the length was only limited by the DMD beam power, that provided the confinement during the cutting procedure. By increasing the DMD power and reducing the disorder, coming from $2D$ lattice, this approach can enable the initialization of even longer chains of controlled length, which are particularly interesting for quantum simulations purposes.
 
 \section{Atom number resolved state readout}
 
 One of the drawbacks of quantum gas microscopes is so called parity projection during the imaging process. When subject to a resonant light two atoms undergo photon assisted collisions, resulting in the molecule formation with a large kinetic energy. The molecules are not trapped by the optical lattice and get lost during this process. This means, that in our image the sites with two particles on the same site appear dark like an empty site and the sites with three particles appear as one. Although it is not a problem in the Mott insulator regime, it leads to the loss of large amounts of information about the state in the superfluid regime. 
 
 However, for a one dimensional system we can utilize the direction transverse to the chain, in order to achieve full number resolution. The idea behind this method is simple: once the dynamics along the chain is frozen (in our case we do this by rapidly increasing the lattice depth along the chain to $45 \textrm{Er}$), we can switch the confinement transverse to the chain (which is provided by the other optical lattice in this case) to spread the atoms from every site of the chain into tubes $\sim 80$ lattice sites long in transverse direction. In this case the probability of doubly occupying any site of the tube is given by
 \begin{equation}
 p(n,L) = \sum_{i=1}^{n-1}\frac{i}{L},
 \end{equation}
 where $n$ is the number of particles that start on the site in the beginning and $L$ is the length of the expansion tube in lattice sites. An example of such process is shown in figure \ref{fig:CTE_fullcount_one_line}.
 
 Although this technique is very powerful for isolated one dimensional systems, we would like to develop a method that would allow us to preform counting of a $2\times N$ plaquettes. At first the solution seems very simple: one just needs a way to isolate the two sides of the plaquette and repeat the above procedure. Using the DMD we can project a single site wide Gaussian beam to disable the tunneling between two sides of the plaquette. However that leads to an issue: the potential, required to completely suppress the hopping between two sides, has a sufficient hight at the position of the atoms, serving as an effective hill the atoms roll down from. This process imparts momentum onto the atom in the direction away from the barrier potential, which is large compared to the diffusion rate of the atoms relative to their center of mass. The result is that atoms can only spread across $\sim 15$ lattice sites, before they leave the field of view of the microscope. Such limited spread results in significant probability of double occupancies, when the initial number of atoms is above $3$, and hence not suited for our purposes.
 
 To mitigate this issue we add another step to this process. After a very short expansion time, when the edge of the atom distribution moves few sites away from the barrier, and it's effect can be neglected. We briefly capture the atoms back into the lattice and then release them again. This process randomizes the momentum of each atom and allows them to expand into corresponding half tube with zero center of mass momentum. This allows us to spread the occupation of a single site of the plaquette into $\sim 50$ lattice sites, resulting in the $94\%$ probability of successfully detecting $3$ atoms or less.
 
  \section{Pure versus mixed states}
In order to understand what does the purity of a state means and why is it important to be able to measure it let's take a closer look at the difference between the quantum superposition and classically mixed states. Although both of them can lead to different measurement outcomes in some bases, there exist a basis in which a pure quantum state will result in only possible measurement outcome. As an example imagine that we have many copies of a spin $\frac{1}{2}$ particle, and we can  measure each of them in along any of the axis $x$, $y$, and $z$. If we had prepared half of the particles in a state $\left| \uparrow \right>$ and the other half in a state $\left| \downarrow \right>$ along the $z$ direction, the measurement in any of the bases will result in the $50/50$ split between the outcomes. Hence we will conclude that the state was mixed. Now, if the particles were initially prepared in the state $(\left| \uparrow  \right>+ \left| \downarrow  \right>)/\sqrt{2}$, the measurement will also lead to the $50/50$ split between the outcomes in $y$ and $z$ basis. However, if we measure this state along $x$ direction, we find that it produces only one of two possible outcomes. 
 
 In terms of formal mathematical notations it is useful to introduce a concept of density matrix. Given a state $\left|\psi \right>$ the density matrix is defined as an outer product of the state with itself, and can be formally written as $\rho_{\mathrm{pure}} = \left|\psi \right> \left< \psi \right|$. It is now easy to generalize this concept to include the mixed states, since they are just a collection of different pure states that occur  with certain probabilities, and can formally be written as $\rho_{\mathrm{mixed}} = \sum_{i} c_{i}\left|\psi_{i} \right> \left< \psi_{i} \right|$, where $c_{i}$ is the probability of state $\left|\psi_{i} \right>$ so that $\sum_{i}c_{i}=1$. This gives us a very intuitive way of quantifying the purity of a given state by looking at the trace of the square of the density matrix. In the case of a pure state 
 \begin{equation}
  Tr(\rho_{\mathrm{pure}}^2) = Tr(\left|\psi \right> \left< \psi \right| \left|\psi \right> \left< \psi \right|) = Tr(\left|\psi \right> \left< \psi \right|) = Tr(\rho_{\mathrm{pure}})=1.
 \end{equation}
 On the other hand, if the state is mixed
 \begin{equation}
 Tr(\rho_{\mathrm{mixed}}^2) = Tr(\sum_{i} c_{i}\left|\psi_{i} \right> \left< \psi_{i} \right| \sum_{j} c_{j}\left|\psi_{j} \right> \left< \psi_{j} \right|) = Tr(\sum_{i} c^2_{i}\left|\psi_{i} \right> \left< \psi_{i} \right|) = \sum_{i} c^2_{i},
 \end{equation}
and since there should be at least two non zero terms in this sum $Tr(\rho_{\mathrm{mixed}}^2) <1$. This means, that by measuring the trace of the density matrix squared, we can tell how pure the state is. 

Now let's consider what happens to the purity of a state, described by a density matrix $\rho$, during the coherent time evolution. The time evolved density matrix can be written as $\rho(t) = U\rho(0) U^+$, where $U$ is the unitary evolution operator, and hence 
\begin{equation}
Tr(\rho^2(t)) = Tr(U\rho(0) U^+ U\rho(0) U^+) = Tr(U\rho(0) \rho(0) U^+) = Tr(U^+ U\rho(0) \rho(0))  = Tr(\rho(0)^2). 
\end{equation}
This means that unitary evolution preserves the purity of the initial state. This also gives us a way to determine whether the dynamics in the system is coherent or not, since any coupling to environment will result in loss of the purity. 

 \section{Hong-Ou-Mandel effect}
In the next few sections we will outline the procedure, that allows us to experimentally measure the purity of a quantum state. A more in depth explanation can be found here \cite{preiss thesis}. In order to understand how this method works, it is instructive to consider how so called Hong-Ou-Mandel effect \cite{HOM} works. The set up is the fallowing: Consider two photons that are incident onto two different input ports of a perfect $50/50$ beamsplitter. Furthermore, let's suppose that they are identical to one an other in every aspect, except for some parameter $\lambda$, which can be anything, from photon frequency to the exact arrival time onto the beamsplitter (for simplicity let's assume that they have different wavelength $\lambda_1$ and $\lambda_2$).

Mathematically the beamsplitter transformation can be written in terms of creation an annihilation operators of the input modes i1, i2 and the output modes o1, o2 as
\begin{equation}
	\begin{aligned}
& a_{i1}^+ \rightarrow (a_{o1}^+ +ia_{o2}^+ )/\sqrt{2} \\
& a_{i2}^+ \rightarrow (ia_{o1}^+ +a_{o2}^+ )/\sqrt{2}.
	\end{aligned}
\end{equation}
When two photons are incident onto two different ports of the beamsplitter, the resulting state can be found as fallowing:
\begin{equation}
\begin{split}
& a_{i1}^+(\lambda_1) a^+_{i2}(\lambda_2) \ket{0} \rightarrow  \frac{1}{2} (a_{o1}^+(\lambda_1) +ia_{o2}^+(\lambda_1) ) (ia_{o1}^+(\lambda_2) +a_{o2}^+(\lambda_2) )\ket{0} = \\
&=\frac{1}{2} (i a_{o1}^+(\lambda_1) a_{o1}^+(\lambda_2) + ia_{o2}^+(\lambda_1)a_{o2}^+(\lambda_1) +a_{o1}^+(\lambda_1) a_{o2}^+(\lambda_2) - a_{o2}^+(\lambda_1) a_{o1}^+(\lambda_2))\ket{0},
\end{split}
\label{eq:BS}
\end{equation}
where $\ket{0}$ corresponds to a vacuum state. We can see, that, in general, the beamsplitter results in $3$ possible states: $\ket{2,0}$, $\ket{0,2}$ and $\ket{1,1}$, where $\ket{m,n}$ represents a state with m photons in the port one and n photons in the port two on the output side of the beamsplitter. However, if the photon infrequences are identical, there is no way to determine the correspondence between the photons on the output port with those at the input, this is where the quantum nature of these particles comes into play. In this case two last terms in the second line of equation~\ref{eq:BS} cancel each other, and the state $\ket{1,1}$ does not occur anymore. This effect can be summarized as following: if two identical photons are incident onto a perfect $50/50$ beamsplitter the probability of simultaneously detecting a photon in each of the output ports equals to zero. 

