%!TEX root = ../dissertation.tex
%\begin{savequote}[75mm]
%Nulla facilisi. In vel sem. Morbi id urna in diam dignissim feugiat. Proin molestie tortor eu velit. Aliquam erat volutpat. Nullam ultrices, diam tempus vulputate egestas, eros pede varius leo.
%\qauthor{Quoteauthor Lastname}
%\end{savequote}

\chapter{Holographic approach to arbitrary potential generation}

\section{Why do we need custom optical potentials?}
The main advantage of cold atoms experiments is an ability to prepare and probe the system on it's characteristic time and length scales (for a lattice models those would be a tunneling time and lattice spacing respectively). Quantum gas microscope enables us to observe our lattice system with a single site resolution, which lead to the first direct observation of superfluid to Mott insulator \cite{Bakr2010, bloch} as well as paramagnet to aniferromagent \cite{Simon2011, Parsons??} quantum phase transitions. However, in order to fully enable the capabilities of such quantum simulators one needs a tool to arbitrary control the shape of optical potentials in the system. This would enable initialization of initial states with high fidelity \cite{us, somebody else?}, studies of the systems with complex \cite{Tilman QPC, Roatti QPC} as well as dynamically changing \cite{??} geometries. To achieve this, one has to control the light intensity on a single lattice site scale, this can be achieved by means of spatial light modulators (SLM).

\section{Spatial light modulators in the image plane}
The first straight forward way is to use the microscope in reverse, by imaging a desired light pattern onto the atoms \cite{P. Schauß thesis, RMA thesis, MAZU thesis}. Although this method is very powerful for creating large scale, smooth-varying potentials, it has two major challenges. \textit{First}, the wavelength of conservative light potentials, that one wants to project, is generally very different from the atom imaging wavelength, which most quantum gas microscopes are optimized for. This can lead to significant distortion of the desired wavefront due to chromatic aberrations, if the imaging system is not specifically corrected for that. First two leading orders of aberrations in this case are given by defocus and spherical aberration. This problem can be solved by either using an achromatic imaging system, that is wavelength insensitive in the desired range, or precompensating the imaged wavefront for the defocus cosed by the imaging. \textit{Second}, it is fundamentally impossible to create $100 \%$ intensity modulation of the potential on the single site scale without using Fourier filtering and phase control of the wave front. 

To illustrate the last point, let's consider a single lens with a focal length $f$ and a fixed aperture of size $d$ in the Fourier plane (see fig~\ref{fig:DMD_lens}). To keep the math simple let's consider only two spatial dimensions $x$ and $z$. The sharpest feature that one can make in the image plane with constant phase wavefront in the Fourier plane is so called point spread function, with the resulting light intensity in the image plane given by $I(x) \sim sinc^2(\frac{\pi}{\Theta} x)$, where $\Theta = \frac{\lambda f}{d}$ (see fig~\ref{fig:DMD_lens} A). The question becomes: how can one make an intensity pattern that is periodic with period $\Theta$? Two different electric field profiles would give us the intensity profile that satisfies this requirement: $E_1 \sim \sqrt{(1+cos(\frac{\pi}{\Theta}x))/2}$ and $E_2 \sim cos(\frac{\pi}{2\Theta} x)$, both yielding $I \sim (1+cos(\frac{\pi}{\Theta}x))/2$ (see fig~\ref{fig:DMD_lens} B,C). The first one has only positive electric field components and hence the same phase, however it's Fourier transform requires the Fourier plane, that is double the size of our initial point spread function. The second one has the same Fourier plane size as the one we started with at the expense of having electric field become negative or alternate the phase between $0$ and $\pi$ between adjacent maxima. This example shows us that using the phase control of the wave front allows us to increase the image plane resolution by a factor of two.
%The characteristic size of the feature is $FWHM = 0.84 \Theta$.

\section{Spatial light modulator in the Fourier plane}
Although, phase control of the electric field using image plane SLM can be achieved using number of phase coding methods along side with spatial filtering in the Fourier plane \cite{Lee1970, Goorden2014}, one still needs to correct for the aberrations, the beam encounters on its way form the SLM to the desired image plane (in our case we call it atom plane). Note, that one needs to correct for both intensity pattern, that eliminates the SLM, as well as any phase deviation from the desired wavefront, which might be particularly challenging for imaging systems with high numerical aperture, due to the breakdown of the paraxial approximation. In order to achieve the best possible quality of the optical potentials it would be desirable to be able to actively measure and compensate for aberration in the system, here the use of the SLM in the Fourier plane becomes particularly useful.

There are two main types of SLMs that are being used in such setups: liquid-crystal display (LED) based and digital micro-mirror device (DMD) based. LEDs can be configured to provide amplitude modulation or phase modulation of the incident beam, using the birefringence of liquid crystals. With optimization and feedback algorithms [83, 84], it is possible to create complex, re-programmable potentials, using this technology. However this type of SLMs have one major drawback when it comes to cold atoms experiments. Due to the nature of liquid crystals, they have to be placed in the into the switching electric filed, that oscillates with frequencies in $kHz$ range, causing the output light to blink at the same frequency. This blinking could potentially be a problem, since the typical trap frequencies of the atoms in optical lattice experiments lay in the same frequency range, and hence the blinking light might induce undesired heating in the system.

Another approach is to use a DMD, that consist of CMOS arrays of micromirrors on torsion hinges which can be switched
between two angles corresponding to the \textit{on} and \textit{off} directions. These devises are typically used in commercial video projectors, and are available with resolutions up to $2560 \times 1600$ pixels. The device used in our experiment is the Keynote Photonics FlexLight X3, which uses the Texas Instruments DLP 5500 chipset. The $1024 \times 768$ micromirror array has a pitch of $10.8 \mathrm{\mu m}$, micromirror tilts $\pm 12^{\circ}$ from normal, and an update rate of $5 \mathrm{kHz}$. The time required to switch between individual patterns is even shorter, on the order of a few $\mathrm{\mu s}$, which can allow dynamical control over the potential.

\section{Encoding phase information with DMD}
Since DMD does not have a build in ways to control the phase of the light field, one needs to find a way to encode this information using binary structure of the DMD pixels. In our experiment we use holographic approach that works as fallowing (see fig.~\ref{fig:DMD_grating}): Consider a plane wave incident onto an amplitude diffraction grating. The grating will crate multiple diffraction orders on the output, each satisfying the condition $k_n = k_0 + n*k_g$, where $k_n$, $k_0$, $k_g \propto \frac{1}{d}$ are the wave vector of the n-th diffraction order, incident light, grating respectively, and $d$ is the period of the grating. We focus on the first diffraction order for the rest of the discussion. The overall spatial translation of the diffraction grating gets mapped onto the phase of the electric field in the diffracted order (see fig~\ref{fig:DMD_grating} B). Also the duty cycle of the grating will result in the intensity modulation of the out going light. Therefore by controlling the phase and duty cycle of the underlying grating we can locally control the phase and amplitude of the out going wave form the DMD \cite{Zupanchich thesis}.

\section{Optical setup}
We chose to use the aperture of $500$ pixels (which approximately equals to $5.4 \mathrm{mm}$) in diameter on our DMD. In order to image it onto the objective Fourier plane, which is $18 mm$ in diameter we use the optical set up shown in the figure~\ref{fig:DMD_setup} A. In order to make the alignment easier and also to be able to control the image plane position of the potential without changing the hologram we use IP and FP mirrors (see fig.~\ref{fig:DMD_setup}). The FP mirror moves the beam primarily in the Fourier plane, since it's located very close to the image plane of the imaging system, and hence does not change the position of the beam in the image plane. The IP mirror position is chosen such that it primarily displaces the beam in the image plane. In order to understand the exact location of this mirror it is instructive to look at where the plane of the DMD (which is the Fourier plane of the overall imaging system with respect to the atoms) is imaged to (see fig~\ref{fig:DMD_setup} B). It is clear from the figure that this mirror is positioned exactly in the Fourier plane of the imaging system and hence moves the beam in the image plane.

\section{Calibration and aberration correction}
In order to achieve diffraction-limited resolution in the atom plane we need a way to map out the aberrations that occur on the optical path from DMD to the atom plane. We also need to compensate for the intensity profile of the beam that illuminates the DMD. 

\section{Holographic potentials and their limits}
To create a desired potential on the atoms we rely on the imaging system property, that electric field in the image plane equals to the Fourier transform of the field in the Fourier plan. Therefore in order to create desired potential we first numerically find it's Fourier transform (note that in general it will be a complex valued function),then in order to represent both amplitude and phase information of the desired potential we use the grating method, that encodes them in local phase and duty cycle of the grating respectively (see fig.~\ref{fig:DMD_hologram}). As long as we know the aberration profile in our system we can easily correct for it by locally modifying the phase to $\phi_{total} = \phi_{target} - \phi_{aberrations}$ and amplitude $A_{total} = \frac{A_{target}}{A_{beam}}$.

If our DMD had pixels with a gray scale, the potential would be limited by the discretization of individual pixels, however the binary nature of the device create an additional noise source that limits the precision of the resulting potentials. One straight forward way to binarize the hologram is to switch pixel with index $(i,j)$ \textit{on} if 
\begin{equation}
\left| k_x i + k_y j + \phi_{total} \; mod \: 2 \pi \right| < arcsin(A_{total}),
\end{equation}
and \textit{off} otherwise. Here $k_x$ and $k_y$ are $x$ and $y$ components of an a priory chosen wave vectors of the underlying grating. Such method has one straightforward problem, when one of the components of the grating wave vector is zero. In this case the stripes of the grating exactly coincide with pixels of the hologram, and if the wave vector is not equal to a integer number of pixels, result in the distortion of the desired potential (see fig.~\ref{fig:DMD_K_rotation} A). This problem can be avoided by slightly randomizing the pixels at the edge of the slit, using the function 
\begin{equation}
\frac{1}{2}(tanh((p(i,j) + arcsin(A_{total}))/\xi) - tanh((p(i,j) - arcsin(A_{total}))/\xi)),
\end{equation}
where $p(i,j))=\left| k_x i + k_y j + \phi_{total} \; mod \: 2 \pi \right|$ and $\xi$ is a parameter that control the "smoothness" of the distribution (see fig.~\ref{fig:DMD_K_rotation} B). Now we compare if the value of this function at any pixel position is grater then a randomly chosen number between $0$ and $1$, we set pixel \textit{on} and \textit{off} otherwise. From figure~\ref{fig:DMD_K_rotation} A it is clear that this method helps to reduce the effect of "bad" grating alignments, however it comes at a cost of increasing overall noise on the potential. This is easy to understand, since the more random we make the potential the more likely it is to flip, otherwise correct pixel value. It fallows form the discussion above, that in order to archive the best performance with the holograms it is beneficial to use nearly deterministic method of binarization away from "bad" angles. 
 
Another parameter that could effect the hologram quality is the absolute value of the grating wave vector, or in simple terms a period of stripes. Although one might expect to find some optimal value for this parameter, our numerical simulations show, that the quality of the resulting potential stays unaffected in a wide range of parameters (see fig.~\ref{fig:DMD_K_vary}A). Surprisingly, even when the stripe width becomes very small and additional grating emerges with separate wave vector (see fig.~\ref{fig:DMD_K_vary}B) the image is not being effected too much. Additionally the unwanted stuff in the image plane can be easily removed by spatial filtering.
 
We also try to determine the optimal value of the parameter $\xi$. To do that, we take the difference between two cuts, that are separated by one lattice site, through the resulting image plane profile. The introduction of the random parameter makes the resulting image plane profile to vary slightly from realization to realization. Therefore one can ask the question, what is the probability to obtain a pattern with desired maximum difference between two cuts? Our numerical results (see fig.~\ref{fig:DMD_rand_vary}) suggest that introduction of a small amount of randomness can lead to the improved results compared to the strictly deterministic case. Although, more complex way to reduce the binazization artifacts like error diffusion algorithms \cite{floyd, Hu2018} have been studied, their application in our system does not improve the results. 

